---
title: "Progress report 1"
author: JB Lim
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

# Parameters
un_voting_file <-
  here::here("GitHub/dcl-2020-01/jb/c01-own/data/un_voting.rds")

country_border_data_democracy_file <-
  here::here("GitHub/dcl-2020-01/jb/c01-own/data/country_border_data_democracy.rds")

path_name <-
  here::here("GitHub/dcl-2020-01/jb/c01-own/data/un_voting_wrangled.rds")

#===============================================================================
# Reading in the data files.
un_voting <-  
  un_voting_file %>%
  read_rds() %>%
  mutate_at(vars(country), ~ str_to_lower(.))

country_info <-
  country_border_data_democracy_file %>%
  read_rds()

```

**Wrangling the UN voting data**

- First, let's explore the data quality of the un voting dataset provided by Professor Erik Voten.

```{r}
un_voting %>%
  summarize_all(~ sum(is.na(.))) %>%
  glimpse()
```

- We notice that "member" has the most number of NAs. As nations joined the UN in different times since its establishment in 1945, any observation with NA values for "member" should be dropped. 

```{r}
un_voting <-
  un_voting %>%
  drop_na(member)
```


- Second, we notice that Czech Republic and Czechoslovakia constitute the nations with member code "9" (not a member). In particular, Czech Republic joined the United Nations in 1993 with the dissolution of Czechoslovakia into Czech and Slovakia. Thus, the data for Czech Republic until 1992 has the member code "9" and that of Czechoslovakia for 1993 onward has member code "9". We would be dropping such data.

```{r}
un_voting %>%
  filter(member == 9) %>%
  glimpse() %>%
  count(countryname, year)

un_voting <-
  un_voting %>%
  filter(member != 9)
```

- Let's also access whether there are any problems with Slovakia.

```{r}
un_voting %>%
  filter(countryname == "Slovakia") %>%
  count(vote)

un_voting %>%
  filter(countryname == "Slovakia", vote == 9) %>%
  distinct(year)
```

- We see that Slovakia had 74 votes under "not a member." Slovakia became a member of the United Nations in 1993. Thus, we see how the votes from 1992 for Slovakia have been counted as "not a member" and instead registered under Czechoslovakia. We drop this value as well.

```{r}
un_voting <-
  un_voting %>%
  filter(countryname != "Slovakia" | vote != 9)
```


- Let's also access in, greater depth, whether other nations face similar issues to Slovakia in terms of their vote choice record. 

```{r}
vote_nine <-
  un_voting %>%
  left_join(
    country_info %>% 
      select(un_member_date, iso),
    by = c("country" = "iso")
  ) %>%
  filter(vote == 9) %>%
  count(countryname, year, un_member_date)

vote_nine %>%
  count(year == un_member_date | year == (un_member_date - 1))

vote_nine %>%
  filter(is.na(un_member_date))

vote_nine %>%
  filter(year != un_member_date, year != (un_member_date - 1))
```

- We notice that most "9" voting codes for most nations come from one or two specific years. Out of 81 such cases 70 (`r 70/81 * 100` %), the year with "9" voting code matched the membership year (or the preceding year). For instance, in the case of Sweden, the nation was admitted in November 19, 1946. Hence, the nation was recorded as "9" for voting in resolutions prior to that date in the same year. We should hence be dropping these values.

- The remaining 11 cases mainly come from nations that dissolved, merged or transferred their membership status since the establishment of the 1945. Indeed, they include China (previously held by Taiwan), Yemen (North Yemen, South Yemen), Zanzibar and Yugoslavia. These "edge" cases will be covered below in greater depth. For now, we drop these value for the similar logic as above. 

```{r}
un_voting <-
  un_voting %>%
  filter(vote != 9)
```


- Finally, aside from Czechoslovakia, we notice that seven countries have been members of the United Nations and currently are not part of the United Nations. The cases can be divided up as below and we check if there are any discrepencies in country names resulting from changing membership status or dissolution of states.


```{r}
# Full list of countries
names <-
  un_voting %>%
  distinct(countryname, country)

names %>%
  anti_join(country_info, by = c("country" = "iso")) %>%
  pull(countryname)
```

1. Taiwan lost its status in the United Nations in 1971. The People's Republic of China replaced Taiwan's seat.

2. Yugoslavia was a member of the United Nations from 1945 to 1992 when it disintegrated into Croatia, Slovenia, Bosnia and Herzegovina, and Serbia and Montenegro (called Federal Republic of Yugoslavia until 2003). Yugoslavia "re-joined" the United Nations in 2000 with the entry of "Serbia and Montenegro" as one federation of state. Starting 2006, the seat was transferred to Serbia after Montenegro's independence from Serbia.

3. Yemen Arab Republic (North Yemen) and Yemen People's Republic (South Yemen) united in 1990 to form the current Republic of Yemen.

4. Similarly, the German Democratic Republic (East Germany) acceded to the Federal Republic of Germany (West Germany) in 1990.

5. Zanzibar, as a semi-autonomous region of Tanzania, joined the UN in 1963, but soon after merged with the United Republic of Tanganyika to form the United Republic of Tanzania on 1964.


Given these information, let's check to see if all the data (including country name corresponding to correct years) are updated and accurate.


```{r}
# Taiwan
un_voting %>%
  filter(countryname == "Taiwan, Province of China", year > 1971) %>% 
  count() %>%
  pull()
```

```{r}
# Zanzibar
un_voting %>%
  filter(countryname == "Zanzibar", year != 1963) %>%
  count() %>%
  pull()
```

- As shown above, the cases for Taiwan (#1) and Zanzibar(#5) are relatively straightforward and the dataset has accurate information for the two states.


```{r}
# Yugoslavia
un_voting %>%
  filter(countryname == "Yugoslavia", year > 1991) %>% 
  count(year)

# Serbia
un_voting %>%
  filter(countryname == "Serbia", year > 1999) %>% 
  count(year, country)

# Montenegro
un_voting %>%
  filter(countryname == "Montenegro", year > 1999) %>% 
  count(year)
```

- For Yugoslavia (#2), we notice how the dataset did not change its name to "Serbia" starting 2000. Specifically, while Montenegro was accurately named starting the year "2006" when it declared independence from Serbia, the data for the name "Serbia" only begins in 2018. Hence, we should rename Yugoslavia starting the year 2000 to "Serbia" to reflect these changes.

```{r}
# rename
un_voting <-
  un_voting %>%
  mutate(
    country = if_else(country == "yug" & year > 1999, "srb", country),
    countryname = 
      if_else(countryname == "Yugoslavia" & year > 1999, "Serbia", countryname)
  )
```

- Let's finally check the data for the Yemens (#3) and Germanys (#4).

```{r}
# Yemen
un_voting %>%
  filter(countryname == "Yemen") %>%
  count(countryname, year)

# Germany
un_voting %>%
  filter(countryname == "Germany") %>%
  count(countryname, year)
```

- We can confirm that the names are merged to their current names staring the 1991 and 1990 respectively.

- We should be aware of these "edge" cases when working with the data in the next part. 


**Note: Tuvalu**

- Tuvalu is a member of the United Nations. However, we were unable to include Tuvalu for "boundaries or information" due to its absence in our original "Natural Earth" dataset.


**Save Answer to rds file**

```{r}
un_voting %>%
  write_rds(path_name)
```

