---
title: "UN Voting Patterns"
author: "JB Lim"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)
# task-begin
library(dcl)
# task-end

# If you will not be providing answers, simply delete the following lines.
# Parameters
# solution-begin
  # File where generated answers are saved, by default in the home directory
file_answers <- "~/answers.rds"
# solution-end
# task-begin
  # File for downloaded answers
file_answers <- ""
# task-end
# solution-begin
  # Save answers
SAVE_ANSWERS <- TRUE
# solution-end

# solution-begin

un_voting_file <-
  here::here("c01-own/data/un_voting_wrangled.rds")

country_information_file <-
  here::here("c01-own/data/country_border_data_democracy.rds")

# Voting Codes
vote_code <-
  c(1, 2, 3)

# Map Aesthetics
MAP_COLORS <- RColorBrewer::brewer.pal(n = 9, name = "PuBu")
MAP_NA_COLOR <- "grey75"

# solution-end


# task-begin

#===============================================================================

# Read in answers
if (str_length(file_answers) > 0) {
  answers <- read_rds(file_answers) 
}
# task-end
```

## Challenge

**Background Context**

- This challenge deals with the **“United Nations General Assembly Voting Data”** posted on the *Harvard Dataverse.* The dataset examines all the “roll-call votes in the UN General Assembly” from 1946 to 2018.

- The code book <https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/LEJUQZ/IPRJ61&version=21.0> of the dataset articulates the variables of the raw UNGA dataset. Notable variables include Vote Choice (with 1 as Yes, 2 as Abstain, 3 as No, 8 as Absent, 9 as Not a member), Country Code, Year of Vote, Votes identified as important by U.S. State Department report and Issue Codes.
  - Issue Codes include: Votes relating to the Palestinian Conflict, Nuclear weapons and nuclear material, Arms control and disarmanent, Colonialism, Human Rights, Economic Development.
  
- The UN voting dataset has already been partially wrangled to account for transitions and dissolutions in certain states such as Yugoslavia or Czechoslovakia. This challenge focuses more on the exploratory data analysis and visualization aspect. 
  
- The country information file contains the boundary and relevant information (from Natural Earth) for 192 UN member states. The information dataset also includes While the United Nations has 193 member states, the information dataset provided excludes Tuvalu due to its absence from the original "Natural Earth" dataset.

- Both datasets can be found in the Data Box <https://stanford.app.box.com/folder/104837039459>.

- Finally, the information dataset includes additional facts on the status of democracy in 2019 (imported from "Freedom in the World") and UN Membership status from the Gapminder dataset covered in the Stanford Data Challenge Lab course. 

**Objective**

In this challenge, we will be examining the status of the United States' hegemonic leadership in the post Cold-War era (1992-2018) by assessing the voting behaviors of other nations in the United Nations General Assembly.

**Reading Data + Wrangling**

__q1__ After reading in both files, filter for UN General Assembly Votes from 1992 to 2018. Then, combine the United Nations voting dataset with the country information dataset. In doing so, note the following:

- We will be dropping "Czechoslovakia" and "Yugoslavia" which ceased to exist in 1992 and 1993 respectively (only represents 77 out of 862969 observations) from the UN General Assembly Dataset.

- Rename:
    "country" to "iso",
    "me" to "palestine",
    "nu" to "nuclear_weapons",
    "di" to "disarmament",
    "hr" to "human_rights",
    "co" to "colonialism",
    "ec" to "economic_dev"
    
- Select the following variables:
  rcid, vote, iso, countryname, year, important vote, me, nu, di, hr, co, ec +
  all the variables in country information.


```{r}
# solution-begin

# UN voting data
un_voting <-  
  un_voting_file %>%
  read_rds() %>%
  mutate(country = str_to_lower(country))

# Country Information
country_info <-
  country_information_file %>%
  read_rds()


# Combining the two data sets
q1 <-
  un_voting %>%
  filter(year > 1991) %>%
  left_join(
    country_info %>% select(-country),
    by = c("country" = "iso")
  ) %>%
  drop_na(electoral_democracy) %>%
  select(
    rcid,
    vote,
    iso = country,
    countryname, 
    year,
    importantvote,
    palestine = me,
    nuclear_weapons = nu,
    disarmament = di,
    human_rights = hr,
    colonialism = co,
    economic_dev = ec,
    pop_est:electoral_democracy
  )
# solution-end

# Print results
if (exists("q1")) q1
# task-begin

# Compare result with answer
if (exists("q1")) compare(answers$q1, q1)
# task-end
```

**Calculating Alignment**

__q2__
Let's try to calculate the "alignment" of votes between the United States and other UN member states for all the unique resolutions since 1992. The rcid code uniquely identifies each distinct resolution from 1992 to 2018. The variable vote indicates the countries vote for the specific resolution. For now, let's focus on the votes "Yes", "No" and "Abstain." Refer to the documentation on the dataset to understand the numeric codes for each corresponding vote.

To do this, we would need the following pieces of information:
1. Retrieve information on the United States' (or any desired nations') voting record for every resolution (rcid).
2. Compare that with other nations through grouping by each rcid.
3. Calculating the difference in votes using absolute values and providing appropriate labels.

__q2.1__ Build a function called "target_nation" that takes in a string of iso code and returns a tibble with two variables: each unique rcid and the target nations' vote.

```{r}
# solution-begin
target_nation <- function(x)
  q1 %>%
  group_by(rcid) %>%
  filter(iso == x) %>%
  select(rcid, vote) %>%
  rename_at(vars(vote), ~ str_c(., "_", x)) %>%
  ungroup()
# solution-end
```

__q2.2__ Build another function called "Alignment" that takes in a tibble of target nation votes and returns a tibble with a new variable "difference." The "difference" variable should be factorized from numeric values into categorical values("Agreed", "Abstained", "Disagreed").

Hint: The variable for the "target nation votes" should have a "general name," so that that the user does not have to manually change the name everytime the alignment is calculated. This may require an additional line of code for the function in quesiton 2.

Hint 2: The target nation being examined should be filtered out before calculating the alignment. 

```{r}
# solution-begin
alignment <- function(df, vars, target_code)
  q1 %>%
  left_join(df, by = "rcid") %>%
  filter(
    iso != target_code,
    vote %in% vote_code,
    {{vars}} %in% vote_code
  ) %>%
  mutate(
    difference = abs(vote - {{vars}}),
    difference = 
      factor(
        difference,
        labels = 
          c("Agreed", "Abstained", "Disagreed")
      )
  )
# solution-end
```

__q3__ Use the two functions above to create a tibble with the voting alignment trend for the United States from 1992 to 2018. Store the answers in q3.

Note: Exclude the column listing the target nation's votes. 

```{r}
# solution-begin
q3 <-
  alignment(target_nation("usa"), vote_usa, "usa") %>%
  select(-vote_usa)
# solution-end

# Print results
if (exists("q3")) q3
# task-begin

# Compare result with answer
if (exists("q3")) compare(answers$q3, q3)
# task-end
```

__q4__ Perform the same task as q3 for China. Store the answers in q4.

```{r}
# solution-begin
q4 <-
  alignment(target_nation("chn"), vote_chn, "chn") %>%
  select(-vote_chn)
# solution-end

# Print results
if (exists("q4")) q4
# task-begin

# Compare result with answer
if (exists("q4")) compare(answers$q4, q4)
# task-end
```

**EDA 1D**

__q5__ Plot the distribution of "difference" from q3 for the United States. Do the same for China using the information from q4. What conclusions can you draw?

# solution-begin

```{r}
difference_bar <- function(df)
  df %>%
  ggplot(aes(difference)) +
  geom_bar(width = 0.5) +
  scale_y_continuous(
    breaks = scales::breaks_width(25000)
  ) +
  theme_minimal()

difference_bar(q3) +
  labs(
    title = 
      "Number of disagreements with U.S. and U.N. member states \nmore than doubles that of agreement",
    x = NULL,
    y = "Count of UNGA votes between 1992-2018"
  )

difference_bar(q4) +
  labs(
    title = 
      "Number of agreements between China and U.N. member states \n was about ten times that of disagreement",
    x = NULL,
    y = "Count of UNGA votes between 1992-2018"
  )
```

- We notice how the number of disagreements with the United States and the United Nations member states more than doubles that of agreement. In the case of China, we observe how the number of agreements between China and U.N. member states was about ten times that of disagreement. Such a stark difference underscores how, among the United Nations member states, China is often viewed more favorably than the United States.

**Important Votes**
- The United Nations dataset has a variable called "importantvote." This variable captures the votes identified as important by the U.S. State Department report on Voting Practices in the United Nations. The important votes for 2018 have not been yet reflected in the dataset and hence been recorded as "NAs".

# solution-end

__q6__ How much percent of the votes in q1 are considered to be "important"? What is the distribution of "difference" for "important" votes in the case of the United States? What conclusions can be drawn?

# solution-begin

```{r}
q6 <-
  q1 %>%
  distinct(rcid, importantvote) %>%
  drop_na(importantvote) %>%
  count(importantvote == TRUE) %>%
  transmute(prop = n / sum(n))

q3 %>%
  filter(importantvote == TRUE) %>%
  ggplot(aes(difference)) +
  geom_bar(width = 0.5) +
  scale_y_continuous(
    breaks = scales::breaks_width(4000)
  ) +
  theme_minimal() +
  labs(
    title = 
      "Number of agreements between U.S. and U.N. member states \n almost matches that of disagreement in important votes",
    x = NULL,
    y =
      "Count of UNGA votes between 1992-2018\n classified as important by the U.S."
  )
```

- Around `r format(100 * q6$prop[2], digits = 2)` percent of the votes were classified to be "important" by the U.S. State Department between 1992 and 2018.

- Surprisingly, for votes classified as "important", nations in agreement with the U.S. almost equalled that of disagreement. This may underscore how the U.S's hegemnoic leadership in international regimes may be enforced through the U.S. securing supportive votes in resolutions that it considers to be vital or important. 

# solution-end

**Visualization**

__q7__ Visualize the "percentage of votes in agreement with the U.S." on a map. Do the same for China. What conclusions can you draw?

Hint: Making a function would simplify the process!

# solution-begin

```{r}
agreement_map <- function(df, important, name)
  df %>%
  count(countryname, iso, difference) %>%
  pivot_wider(
    names_from = difference,
    values_from = n
  ) %>%
  rename_all(str_to_lower) %>%
  group_by(countryname, iso) %>%
  summarize(prop_agree = agreed / sum(agreed, abstained, disagreed)) %>%
  ungroup() %>%
  full_join(country_info %>% select(iso, geometry), by = "iso") %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = prop_agree), size = 0.01) +
  guides(
    fill = guide_colorbar(
      barheight = 0.5,
      barwidth = 15,
      title = NULL
    )
  ) +
  theme_void() +
  theme(
    legend.direction = "horizontal",
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  labs(
    subtitle = 
      str_c("Percentage of ", important,"votes in agreement with ", name, " (1992-2018)"),
    caption = "Source: UNGA Voting Database"
  )
```

```{r}
# Map for the United States
agreement_map(q3, "", "the U.S.") +
  scale_fill_gradientn(
    breaks = seq(0.1, 0.7, 0.1),
    labels = 
      c("10-%", "20%", "30%", "40%", "50%",
        "60%", "70+%"),
    colors = MAP_COLORS,
    na.value = MAP_NA_COLOR
  ) 
```

- The map above indicates how only Israel, Canada, Australia, France and the UK were in relatively strong agreement with the United States in the United Nations General Assembly. Traditional United States' allies including Korea, Japan, New Zealand, and European nations were positioned in the "middle ground" in terms of their voting alignment with the United States.

- The rest of the world had low voting alignment percentage with the United States. 

```{r}
# Map for China
agreement_map(q4, "", "China") +
  scale_fill_gradientn(
    breaks = seq(0.2, 0.9, 0.1),
    labels = 
      c("20%-", "30%", "40%", "50%",
        "60%", "70%", "80%", "90%+"),
    colors = MAP_COLORS,
    na.value = MAP_NA_COLOR
  )
```

- The map above indicates how only United States and Israel were in strong disagreement with China in their United Nations General Assembly voting pattern. Countries in Africa, Middle East, South East Asia, Latin Americ and Central Asia (alongside North Korea) strongly agreed with China. Some even displayed over a 90% alignment in votes.

- Traditional United States' allies including Korea, Japan, New Zealand, and Western European nations also had relatively "strong" alignment percentages with China.

# solution-end

__q8__ Visualize the "percentage of votes in agreement with the U.S." for votes classified as important by the State Department on a map. Do the same for China. What conclusions can you draw?

# solution-begin

```{r}
agreement_map(q3 %>% filter(importantvote == TRUE), "important ", "the U.S.") +
  scale_fill_gradientn(
    breaks = seq(0.1, 0.8, 0.1),
    labels = 
      c("10%-", "20%", "30%", "40%", "50%",
        "60%", "70%", "80%+"),
    colors = MAP_COLORS,
    na.value = MAP_NA_COLOR
  )
```

```{r}
agreement_map(q4 %>% filter(importantvote == TRUE), "important ", "China") +
  scale_fill_gradientn(
    breaks = seq(0.2, 0.8, 0.1),
    labels = 
      c("20%-", "30%", "40%", "50%",
        "60%", "70%", "80%+"),
    colors = MAP_COLORS,
    na.value = MAP_NA_COLOR
  )
```


- The maps above indicates how, for votes classified as "important" by the U.S. State Department, nations mostly increased in their voting alignment with the United States. Canada, Australia and Israel turned "dark blue" indicating their strong agreement in voting. United States' allies such as Korea, Japan, and European nations also increased in their alignment percentage with the United States. Countries in South America, Africa and Central Asia also increased in their alignment with the United States. In the mean time, China remained to be in strong disagreement with the United States. 

- Such an increase in voting alignment further underlines how, while the nations may disagree with the United States, realist and constructivist alliance considerations compel them to vote in alignment with the hegemon on key resolutions.


```{r}
general <-
  q3 %>%
  count(countryname, iso, difference) %>%
  pivot_wider(
    names_from = difference,
    values_from = n
  ) %>%
  rename_all(str_to_lower) %>%
  group_by(countryname, iso) %>%
  summarize(prop_agree = agreed / sum(agreed, abstained, disagreed)) %>%
  ungroup()

important <-
  q3 %>%
  filter(importantvote == TRUE) %>%
  count(countryname, iso, difference) %>%
  pivot_wider(
    names_from = difference,
    values_from = n
  ) %>%
  rename_all(str_to_lower) %>%
  group_by(countryname, iso) %>%
  summarize(
    prop_agree_important = agreed / sum(agreed, abstained, disagreed)
  ) %>%
  ungroup()

general %>%
  left_join(
    important %>%  select(iso, prop_agree_important),
    by = "iso"
  ) %>%
  transmute(
    countryname,
    prop_increase = (prop_agree_important - prop_agree) / prop_agree
  ) %>%
  arrange(desc(prop_increase)) %>%
  top_n(10)


general %>%
  left_join(
    important %>%  select(iso, prop_agree_important),
    by = "iso"
  ) %>%
  transmute(
    countryname,
    prop_increase = (prop_agree_important - prop_agree) / prop_agree
  ) %>%
  arrange(desc(prop_increase)) %>%
  top_n(-10)
```

- We notice how except for four nations (China, Syria, Cuba and North Korea), all nations increased in their agreement percentage with the United States as the votes became classified as "important" by the State Department. 

# solution-end

**U.S. and China Differs** (Optional)

__q9__ Consider the resolutions where the U.S. and China had opposing votes. 

Which voting option did each nation "lean most favorably towards" over 1992-2018? Was there any difference for votes classified as "important"?

# solution-begin

```{r}
us_china_diff <-
  q3 %>%
  filter(iso == "chn", difference == "Disagreed") %>%
  #  filter_at(vars(me, nu, di, co, hr, ec), all_vars(. == FALSE)) %>%
  pull(rcid)

us_china_map <- function(df)
  df %>%    
  filter(
    rcid %in% us_china_diff, 
    iso != "chn"
  ) %>%
  count(iso, difference) %>%
  pivot_wider(
    names_from = difference,
    values_from = n
  ) %>%
  rename_all(str_to_lower) %>%
  group_by(iso) %>%
  mutate(
    America = replace_na(agreed, 0),
    China = replace_na(disagreed, 0),
    Abstained = replace_na(abstained, 0),
    sum = sum(America, Abstained, China)
  ) %>%
  summarize_at(vars(America, Abstained, China), ~ . / sum) %>%
  group_by(iso) %>%
  mutate(
    top_prop = 
      c("America", "Abstained", "China")
        [which.max(c(America, Abstained, China))]
  ) %>%
  ungroup()




country_info %>%
  left_join(us_china_map(q3), by = "iso") %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = top_prop), size = 0.01) +
  theme_void() +
  labs(
    title =
      "Most voted option when U.S. and China had opposing votes (1992-2018)",
    fill = NULL
  )

```

```{r}
country_info %>%
  left_join(
    us_china_map(q3 %>% filter(importantvote == TRUE)), by = "iso"
  ) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = top_prop), size = 0.01) +
  theme_void() +
  labs(
    title =
      "Most voted option when U.S. and China had opposing votes (1992-2018)",
    subtitle = "Votes that the U.S. State Department classified as important",
    fill = NULL
  )

```

# solution-end

__q10__ Finally, the democracy and income variables on to the map. Are there any possible correlations?

# solution-begin

```{r}
country_info %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = electoral_democracy), size = 0.01) +
  theme_void() +
  labs(
    fill = "Electoral Democracy?",
    subtitle = "Map of Electoral Democracies as of 2019",
    caption = "Source: Freedom House"
  )

country_info %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = income_grp), size = 0.01) +
  theme_void() +
  labs(
    fill = "Income Level",
    subtitle = "Map of Nation's Income Levels",
    caption = "Source: Gapminder Data"
  )

```


- Democratic and high-income nations (Canada, Europe, Australia, New Zealand, Israel, Korea, Japan) were likely to vote in alignment with the United States. 

- Non-democracies were likely to vote in alignment with China. So were middle income nations such as Brazil, India South Africa, Indonesia, Mexico, and Egypt. 

# solution-end

<!-- solution-begin -->
## Answers

To create an RDS file with answers, save all of your solutions in variables such as `q1`, `q2.1`, etc. The following code will create an answer file when you knit the solution.Rmd file. You specify where the answer file is saved using the `file_answers` variable in the parameters.

To provide answers, set `eval=TRUE` in the chunk below.
If you will not be providing answers, simply delete the following lines.

Save answers.

```{r, eval=FALSE}
if (SAVE_ANSWERS) {
  ls(pattern = "^q[1-9][0-9]*(\\.[1-9][0-9]*)*$") %>%
    str_sort(numeric = TRUE) %>% 
    set_names() %>% 
    map(get) %>%
    discard(is.ggplot) %>%
    write_rds(file_answers)
}
```
<!-- solution-end -->
